---
title: "Letters"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Supplemental R code to "Exploratory inference: localizing relevant effects with confidence"
by A. Solari and J. Goeman. 

## Introduction

Suppose we have a moderate-scale spatial problem, structured in a $m = 100 × 100$ square matrix, where for each unit (pixel) we measure the local magnitude of the effect. 

Assume a random sample of size $n=20$ from $m$-variate Gaussian
 $$\mathcal{N}_m(\mu, \Sigma)$$
 with $\mu=(\mu_1,\ldots,\mu_m)^\mathsf{T}$ and $\mathrm{diag}(\Sigma) = (\sigma^2_1,\ldots,\sigma_m^2)^\mathsf{T}$. 

Load the $n\times m$ data matrix:

```{r, message=F, warning=F, error=F, comment=NA}
library(repmis)
source_data("https://github.com/aldosolari/letters/blob/main/rawdata.RData?raw=true")
x[1:10, 1:5]
```

 
The parameter of interest is the standardized effect
$$
 \theta_i = \frac{\mu_i}{\sigma_i}
$$




```{r, message=F, warning=F, error=F, comment=NA}
library(repmis)
source_data("https://github.com/aldosolari/letters/blob/main/letters.RData?raw=true")
source_data("https://github.com/aldosolari/letters/blob/main/rawdata.RData?raw=true")
ls()
```

